### 2024-09-11 appendix_exp1.Rmd
before final edits of revisions round 1

>>>>

## Minor deviation from preregistration

```{r}
# T-test
t_test_accuracy <- t.test(confidence ~ convergence, data = exp1, paired = TRUE) %>% clean_t_test()
t_test_competence <- t.test(competence ~ convergence, data = exp1, paired = TRUE) %>% clean_t_test()
```

In our preregistration, we stated we would run paired t-tests and only conduct mixed models as a robustness check. However, it became clear to us later that mixed models are more appropriate in light of our data structure. In the main article, we therefore only report the results of the mixed models. Note, however, that using paired t-tests, we confirm our results (Accuracy: `r t_test_accuracy$estimate`, t-statistic = `r t_test_accuracy$statistic`, p `r t_test_accuracy$p.value`; Competence: `r t_test_competence$estimate`, t-statistic = `r t_test_competence$statistic`, p `r t_test_competence$p.value`)

## Sensitivity simulation

We had based our power calculations on a paired t-test (using the widely used tool G*Power), which does not correspond to the mixed-models we eventually ran. To address this point, here, we seek to identify the minimum effect size that our analysis with mixed models could have detected, given a 90% power, an N of 200 participants, and other parameters derived from the estimates observed in the sample of Experiment 1 (see \@ref(tab:exp1-regression-table)). By contrast to a standard power simulation (which varies sample size), a sensitivity simulation varies the effect size, while holding the sample size and desired power constant.

```{r exp1-regression-table}
modelsummary::modelsummary(list("Accuracy" = exp1_model_accuracy, 
                                "Competence" = exp1_model_competence), 
                           title = "Model Results Experiment 1")
```

```{r}
# Sensitivity analysis Experiment 1

library(tidyverse)
library(purrr)
library(broom)
library(lmerTest)

set.seed(1000)

# Generate data
create_data <- function(n_subj, n_div, n_conv, beta_0, beta_1, 
                        tau_0, tau_1, rho, sigma) {

  items <- data.frame(
    item_id = seq_len(n_div + n_conv),
    convergence = rep(c("div", "conv"), c(n_div, n_conv)),
    # make the same as a numeric variable to generate competence
    convergence_num = rep(c(-0.5, 0.5), c(n_div, n_conv)))

  # variance-covariance matrix
  cov_mx  <- matrix(
    c(tau_0^2,             rho * tau_0 * tau_1,
      rho * tau_0 * tau_1, tau_1^2            ),
    nrow = 2, byrow = TRUE)

  subjects <- data.frame(subj_id = seq_len(n_subj),
                         MASS::mvrnorm(n = n_subj,
                                       mu = c(T_0s = 0, T_1s = 0),
                                       Sigma = cov_mx))

  crossing(subjects, items)  %>%
    mutate(
           # turn convergence into factor variable
           # for later analysis. Make sure the levels correspond to the 
           # numeric variables we generate competence with
           convergence = as.factor(convergence),
           convergence = fct_relevel(convergence, "div", "conv"),
           # residual error
           e_si = rnorm(nrow(.), mean = 0, sd = sigma),
           # generate competence
           competence = beta_0 + T_0s + (beta_1 + T_1s) * convergence_num + e_si) %>%
    select(subj_id, item_id, convergence, competence, convergence_num)
}

# test
# data <- create_data(
#   n_subj = 100,
#   n_div  =  2,   # number of low convergence stimuli
#   n_conv =  2,   # number of high convergence stimuli
#   beta_0 = 4,
#   beta_1 = 0.5,  # effect of convergence
#   tau_0      = 0.7,   # by-subject random intercept sd
#   tau_1      =  0.9,   # by-subject random slope sd
#   rho        = 0.1,   # correlation between intercept and slope by-subject
#   sigma      = 0.99) # residual (standard deviation)

# Our estimation function
est_model <- function(n_subj, n_div, n_conv, beta_0, beta_1,
                      tau_0, tau_1, rho, sigma) {

  data_sim <- create_data(n_subj, n_div, n_conv, beta_0, beta_1,
                          tau_0, tau_1, rho, sigma)
  model_sim <- lmer(competence ~ convergence + (1 + convergence | subj_id),
                    data_sim)

  tidied <- broom.mixed::tidy(model_sim)
  sig <- tidied$p.value[2] < .05

  return(sig)
}

# test
# model_result <- est_model(
#   n_subj = 100,
#   n_div  =  2,   # number of low convergence stimuli
#   n_conv =  2,   # number of high convergence stimuli
#   beta_0 = 4,    # intercept (average of divergent condition)
#   beta_1 = 0.5,  # effect of convergence
#   tau_0      = 0.7,   # by-subject random intercept sd
#   tau_1      =  0.9,   # by-subject random slope sd
#   rho        = 0.1,   # correlation between intercept and slope by-subject
#   sigma      = 0.99) # residual (standard deviation)

# Iteration function
iterate <- function(n_subj, n_div, n_conv, beta_0, beta_1,
                    tau_0, tau_1, rho, sigma, iterations) 
{ # number of iterations

  results <-  1:iterations %>%
    map_dbl(function(x) {
      # To keep track of progress
      if (x %% 100 == 0) {print(paste("iteration number ", x))}

      # Run our model and return the result
      return(est_model(n_subj, n_div, n_conv, beta_0, beta_1,
                       tau_0, tau_1, rho, sigma))
    })

  # We want to know statistical power, 
  # i.e., the proportion of significant results
  return(mean(results))
}

# test
# power <- iterate(
#   n_subj = 100,
#   n_div  =  2,   # number of low convergence stimuli
#   n_conv =  2,   # number of high convergence stimuli
#   beta_0 = 4, 
#   beta_1 = 0.5,  
#   tau_0      = 0.7,   # by-subject random intercept sd
#   tau_1      =  1,   # by-subject random slope sd
#   rho        = 0.1,   # correlation between intercept and slope by-subject
#   sigma      = 1,  # residual (standard deviation)
#   iterations = 500)

sensitivity_analysis <- function(file_name, effect_sizes, ...) {

  # only run analysis if a file with that name does not yet exists
  if (!file.exists(paste0("Experiment_1/data/", file_name))) {

    # do the `iterate()` function for each effect size and store the results
    # in a common data frame
    power <- purrr::map_df(effect_sizes, 
              function(beta_1){
                # this is essentially a for loop - 
                # do the following for each 
                # element of effect_sizes

                # To keep track of progress
                print(paste("tested effect size = ", beta_1))

                # run power calculation
                power <- iterate(beta_1 = beta_1, ...)

                # make data frame
                result <- tibble(effect_size = beta_1, 
                                 power = power)

                return(result)

              })

    # write out file
    write_csv(power, paste0("Experiment_1/data/", file_name))
  }
}

# test (note that this requires building the parameter lists below before)
# effect_sizes <- c(1, 1.2)
# test <- purrr::map_df(effect_sizes, 
#               function(beta_1){
#                 # this is essentially a for loop - 
#                 # do the following for each 
#                 # element of effect_sizes
#                 
#                 # To keep track of progress
#                 print(paste("tested effect size = ", beta_1))
#                 
#                 # run power calculation
#                 power <- do.call(iterate, c(parameters_accuracy, 
#                                              beta_1 = beta_1))
#                 
#                 # make data frame
#                 result <- tibble(effect_size = beta_1, 
#                                  power = power)
#                 
#                 return(result)
#                 
#               })
```

```{r}
# Define parameters

# Extract parameters from model estimates of Experiment 1
estimates_accuracy <- exp1_model_accuracy %>% 
  tidy() %>% 
  split(.$term)

estimates_competence <- exp1_model_competence %>% 
  tidy() %>% 
  split(.$term)

# Set parameter lists
parameters_accuracy <- list(
  n_subj = 200,
  n_div  =  2,   # number of low convergence stimuli
  n_conv =  2,   # number of high convergence stimuli
  beta_0 = estimates_accuracy$'(Intercept)'$estimate, # intercept (mean of divergent)
  #beta_1 = estimates_accuracy$convergenceconvergent$estimate, # mean difference convergence vs divergence
  tau_0      = estimates_accuracy$'sd__(Intercept)'$estimate,   # by-subject random intercept sd
  tau_1      =  estimates_accuracy$sd__convergenceconvergent$estimate,   # by-subject random slope sd
  rho        = estimates_accuracy$'cor__(Intercept).convergenceconvergent'$estimate,   # correlation between intercept and slope by-subject
  sigma      = estimates_accuracy$sd__Observation$estimate,  # residual (standard deviation)
  iterations = 1000)

# Set parameter lists
parameters_competence <- list(
  n_subj = 200,
  n_div  =  2,   # number of low convergence stimuli
  n_conv =  2,   # number of high convergence stimuli
  beta_0 = estimates_competence$'(Intercept)'$estimate, # intercept (mean of divergent)
  # beta_1 = estimates_competence$convergenceconvergent$estimate, # mean difference convergence vs divergence
  tau_0      = estimates_competence$'sd__(Intercept)'$estimate,   # by-subject random intercept sd
  tau_1      =  estimates_competence$sd__convergenceconvergent$estimate,   # by-subject random slope sd
  rho        = estimates_competence$'cor__(Intercept).convergenceconvergent'$estimate,   # correlation between intercept and slope by-subject
  sigma      = estimates_competence$sd__Observation$estimate,  # residual (standard deviation)
  iterations = 1000)
```

```{r}
# Run analysis

# Accuracy
do.call(sensitivity_analysis, c(parameters_accuracy, 
                                        list(file_name = "sensitivity_accuracy.csv", 
                                             effect_sizes = c(0.1, 0.2, 0.4, 0.6, 0.8, 1, estimates_accuracy$convergenceconvergent$estimate)
                                             )
                                        )
                )

sensitivity_accuracy <- read_csv("Experiment_1/data/sensitivity_accuracy.csv")

# Competence
do.call(sensitivity_analysis, c(parameters_competence, 
                                        list(file_name = "sensitivity_competence.csv", 
                                             effect_sizes = c(0.1, 0.2, 0.4, 0.6, 0.8, 1, estimates_competence$convergenceconvergent$estimate)
                                             )
                                        )
                )

sensitivity_competence <- read_csv("Experiment_1/data/sensitivity_competence.csv")

```
(ref:sensitivity) Results of the sensitivity analysis.

```{r sensitivity, fig.cap="(ref:sensitivity)"}
# Plot

# combine data
plot_data <- bind_rows(sensitivity_accuracy %>% 
                         mutate(outcome = "Accuracy"), 
                       sensitivity_competence %>% 
                         mutate(outcome = "Competence"))

# plot results
plot_power <- function(data) {

  ggplot(data, 
       aes(x = effect_size, y = power)) +
  geom_point(color = "red", size = 1) +
  geom_line(size = 1, alpha = 0.5) + 
  # add a horizontal line at 90%, our power_threshold
  geom_hline(aes(yintercept = .9), linetype = 'dashed') + 
  # display iterations
  annotate("text", x = 1, y = 0.25, label = "iterations per \n effect size = 1'000")+
  # Prettify!
  theme_minimal() + 
    scale_colour_viridis_d(option = "plasma") + 
  scale_y_continuous(labels = scales::percent, breaks = ) + 
  labs(x = 'Effect size', y = 'Power', 
       title = "Sensitivity Analysis",
       subtitle = "(Finding the minimum detectable effect)") +
    facet_wrap(~outcome)
}

plot_power(plot_data)
```

The results of this can be seen in Fig. \@ref(fig:sensitivity). In our sample, we found a statistically significant effect size of `r estimates_accuracy$convergenceconvergent$estimate` for accuracy and an effect size of `r estimates_competence$convergenceconvergent$estimate` for competence (both on their original scales from 1 to 7). The simulations show that, given our parameter assumptions (based on the observations in our sample) and a sample size of N = 200, we would have detected an effect as small as 0.4 with a power of at least 90% (in fact nearly 100%) for both accuracy and competence.
